A molecular dynamics study of the solvation of carbon dioxide and other compounds in the ionic liquids \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]}

### 2018-09-16 ionic_liquids.tex
Revision of Dilute CO2 solutions section

\usepackage[colorlinks,citecolor=blue]{hyperref}
%\usepackage{lineno}
%\modulolinenumbers[5]
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[]{natbib}
\usepackage[version=4]{mhchem}
\usepackage{adjustbox}   
\usepackage[position=top,captionskip=-12pt]{subfig}
\usepackage{graphicx}
\DeclareGraphicsExtensions{.pdf,.eps,.png}

@@

In the case of ionic liquid \ce{[emim][NTf_2]}, we use the model of K\"{o}ddermann \textit{et al}. \cite{K_ddermann_2007}, which was also derived by refining parameters of the force field by P{\'{a}}dua and Canongia Lopes \cite{Canongia_Lopes_2006}.
The authors characterized this refinement as unusual, given that they modified the LJ parameters while keeping the partial charges unchanged.
Kerl \textit{et al}. \cite{Kerl__2017} recently employed this model to accurately predict Henry constants of \ce{CO_2} in \ce{[emim][NTf_2]} at several temperatures.

@@

\begin{figure*}
\centering
\subfloat[][Koller~\textit{et al}.~\cite{Koller_2012}]{
  \includegraphics[width=.4\linewidth]{koller}}
\subfloat[Batista~\textit{et al}.~\cite{Batista_2015}]{  
  \includegraphics[width=.4\linewidth]{batista}}

\subfloat[Liu~\textit{et al}.~\cite{Liu_2014}]{%
  \includegraphics[width=.4\linewidth]{Liu}}%
\subfloat[Weber and Kirchner~\cite{Weber_2016}]{%
  \includegraphics[width=.4\linewidth]{Weber}}%
\caption{Combined distribution functions of the dihedral angles corresponding to atoms C1-N2-C5-C6 and C2-N2-C5-C6 in the cation \ce{[emim]^+} (see Fig.~\ref{fig:atoms_id}), as sampled in NPT-MD simulations of 250 ion pairs using each model for \ce{[emim][B(CN)_4]}.}
\label{fig:dihedrals_emim}
\end{figure*}

@@

A study of conformations was also carried out for \ce{[emim][NTf_2]} using the model of K\"{o}ddermann \textit{et al}. \cite{K_ddermann_2007}.
A CDF obtained for the dihedral angles of sequences C1-N2-C5-C6 and C2-N2-C5-C6 in \ce{[emim]^+}, which is made available in the Supplementary Material, resembles the one in Fig.~\ref{fig:dihedrals_emim}(d).
The resemblance was expected because the intramolecular parameters employed for \ce{[emim]^+} are identical in these two cases, since the corresponding models are both based on the force field developed by Canongia-Lopes and P\'{a}dua \cite{Canongia_Lopes_2006}.
However, the most likely value of $\psi_1$ is shifted to $0^{\circ}$ in the case of \ce{[emim][NTf_2]}.
This highlights the fact that altering atomic charges can induce conformational changes and should, therefore, be accompanied by a revision of torsion parameters in some cases.
With regard to the anion \ce{[NTf_2]^-}, Deetlefs \textit{et al}. \cite{Deetlefs_2006} experimentally determined the conformer population of the liquid-phase and observed that \ce{[NTf_2]^-} is mostly found in its \textit{trans} conformation.
In Fig.~\ref{fig:die_ntf2}, we present the CDF obtained for the contiguous sequences C1-S1-N1-S2 and S1-N1-S2-C2 in \ce{[NTf_2]^-}, which qualitatively confirms the observation of Deetlefs \textit{et al}. \cite{Deetlefs_2006}.
This CDF also makes clear that \ce{[NTf_2]^-} exhibits a considerable degree of flexibility.
Due to this observed behaviors of both ions, all simulations of \ce{[emim][NTf_2]} in the present work are performed considering flexible cations and anions, as well as DSF electrostatics.
This approach results in a mean density of $1.495 \pm 0.009 ~\mathrm{g/cm^3}$ at $T = 298.0~\mathrm{K}$ and $P = 1.0~\mathrm{bar}$.
The relative error of this result with respect to the experimental value obtained by Tokuda \textit{et al}. \cite{Tokuda_2005} is only 1.12\%.

\begin{figure}
\centering
\includegraphics[width=\linewidth]{Ludwig_anion}
\caption{Combined distribution function of the dihedral angles of the contiguous chains C1-S1-N1-S2 and S1-N1-S2-C2 of anion \ce{[NTf_2]^-}, obtained from NPT-MD simulations of 250 ion pairs of \ce{[emim][NTf_2]} using the model of K\"{o}ddermann \textit{et al}. \cite{K_ddermann_2007}.}
\label{fig:die_ntf2}
\end{figure}

@@

\begin{table*}
	\centering
	\caption{Results of solvation free energies and Henry's law constants of \ce{CO_2} in \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]} at $T = 298.15~\mathrm{K}$ and $P = 1~\mathrm{atm}$. The experimental Henry constants of \ce{CO_2} in \ce{[emim][B(CN)_4]} and in \ce{[emim][NTf_2]} are, respectively, $38.9 \pm 0.03 ~\mathrm{atm}$ \cite{Mahurin_2010} and $39.0 \pm 0.1 ~\mathrm{atm}$ \cite{Finotello_2008}.}
	\begin{adjustbox}{max width=\textwidth}
		\begin{tabular}{ccccccc}  
			\hline\hline
			Model & $\tilde{\rho}_\text{sim}$ & $\Delta G_\text{vdW}$  & $\Delta G_\text{Coul}$  & $\Delta G_\text{sim}$ & $K_s$ & error \\
			& (mol/dm\textsuperscript{3}) & (kcal/mol) & (kcal/mol) &  (kcal/mol) & (atm)  & (\%) \\
			\hline
%			\multicolumn{7}{c}{\ce{[emim][B(CN)_4]}: $K_{s}^{\text{exp}} = 38.9 \pm 0.03 ~\mathrm{atm}$ \cite{Mahurin_2010}} \\
			\multicolumn{7}{c}{\ce{[emim][B(CN)_4]}} \\
			Koller \textit{et al}. \cite{Koller_2012} & $4.5925$ & $0.45 \pm 0.01$ & $-1.06 \pm 0.02$ & $-0.61 \pm 0.02$ & $40 \pm 1$ & $2.8$ \\
			Batista \textit{et al}. \cite{Batista_2015} & $4.3700$ & $0.378 \pm 0.008$ & $-1.24 \pm 0.01$  & $-0.86 \pm 0.01$ & $24.9 \pm 0.5$ & $38.9$ \\
			Liu \textit{et al}. \cite{Liu_2014} & $4.3815$ & $0.226 \pm 0.009$ & $-1.11 \pm 0.01$ & $-0.88 \pm 0.01$ & $24.3 \pm 0.5$ & $37.6$  \\
			Weber and Kirchner \cite{Weber_2016} & $4.5221$ & $0.478 \pm 0.009$ & $-1.23 \pm 0.02$ & $-0.75 \pm 0.02$ & $31 \pm 1$ & $20.3$  \\
			\hline
%			\multicolumn{7}{c}{\ce{[emim][NTf_2]}: $K_{s}^{\text{exp}} = 39.0 \pm 0.1 ~\mathrm{atm}$ \cite{Finotello_2008} }\\
			\multicolumn{7}{c}{\ce{[emim][NTf_2]}} \\
			K\"{o}ddermann \textit{et al}. \cite{K_ddermann_2007} & $3.8205$ & $0.389 \pm 0.005$ & $-0.901 \pm 0.006$ & $-0.512 \pm 0.008$ & $39.4 \pm 0.5$  & $0.97$  \\
			\hline\hline
			\label{table:henry} 
		\end{tabular}
	\end{adjustbox}
\end{table*}

@@

The $\Delta G_\text{sim}$ values for the models of Koller \textit{et al}. \cite{Koller_2012} and K\"{o}ddermann \textit{et al}. \cite{K_ddermann_2007} suggest a more favorable solvation of \ce{CO_2} in \ce{[emim][B(CN)_4]} than in \ce{[emim][NTf_2]}. 
The spatial distribution functions (SDFs) \cite{Svishchev_1993} shown in Figs.~\ref{fig:sdf_ions} confirm this result.
The SDFs of the ions around a \ce{CO_2} molecule were obtained by using the software packages TRAVIS \cite{Brehm_2011} and VMD \cite{HUMP96}.
For that, we carried out NPT-MD simulations of \ce{CO_2}-\ce{[emim][B(CN)_4]} and \ce{CO_2}-\ce{[emim][NTf_2]} systems with a \ce{CO_2} molar fraction of 0.1.
The isovalues were visually determined so as to observe the first solvation shell.
If we compare Figs.~\ref{fig:sdf_ions} (a) and (b), one notes that the region with a low probability of finding an ion is larger for \ce{[emim][NTf_2]}.
Also, the first solvation shell due to \ce{[B(CN)_4^-} occurs at a slightly shorter distance than for \ce{[NTf_2]^-}, as one sees from Fig.~\ref{fig:sdf_ions}(c).

@@

\begin{table}
	\centering
	\caption{Thermodynamic conditions employed in the solvation study at high concentrations of \ce{CO_2} in \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]}~\cite{Makino_2014,Schilderman_2007}.}
	\begin{tabular}{ c  c  c  }  
		\hline\hline
		IL & \ce{[emim][B(CN)_4]}  & \ce{[emim][NTf_2]} \\
		\hline
		model &  Ref.~\citenum{Koller_2012} & Ref.~\citenum{K_ddermann_2007} \\
		T  & 313.15 K & 312.13 K \\
		P  & 21 atm & 37 atm \\
		$x_s$ & 0.352 & 0.479 \\
		$m_s$ & 2.4 mol/kg & 2.35 mol/kg \\
		\hline\hline
		\label{table:solv} 
	\end{tabular}
\end{table}

@@

\bibliography{mybibfile}

>>>>

\usepackage[colorlinks,citecolor=blue]{hyperref}
%\usepackage{lineno}
%\modulolinenumbers[5]
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[]{natbib}
\usepackage[version=4]{mhchem}
\usepackage{adjustbox}   
\usepackage[position=top,captionskip=0pt]{subfig}
\usepackage{graphicx}
\DeclareGraphicsExtensions{.pdf,.eps,.png}

@@

In the case of ionic liquid \ce{[emim][NTf_2]}, we use the model of K\"{o}ddermann \textit{et al}. \cite{Koddermann_2007}, which was also derived by refining parameters of the force field by P{\'{a}}dua and Canongia Lopes \cite{Canongia_Lopes_2006}.
The authors characterized this refinement as unusual, given that they modified the LJ parameters while keeping the partial charges unchanged.
Kerl \textit{et al}. \cite{Kerl__2017} recently employed this model to accurately predict Henry constants of \ce{CO_2} in \ce{[emim][NTf_2]} at several temperatures.

@@

\begin{figure*}
\centering
\subfloat[][Koller~\textit{et al}.~\cite{Koller_2012}]{
  \includegraphics[width=.4\linewidth,trim={0 0 0 12pt},clip]{koller}}
\subfloat[Batista~\textit{et al}.~\cite{Batista_2015}]{  
  \includegraphics[width=.4\linewidth,trim={0 0 0 12pt},clip]{batista}}

\subfloat[Liu~\textit{et al}.~\cite{Liu_2014}]{%
  \includegraphics[width=.4\linewidth,trim={0 0 0 12pt},clip]{Liu}}%
\subfloat[Weber and Kirchner~\cite{Weber_2016}]{%
  \includegraphics[width=.4\linewidth,trim={0 0 0 12pt},clip]{Weber}}%
\caption{Combined distribution functions of the dihedral angles corresponding to atoms C1-N2-C5-C6 and C2-N2-C5-C6 in the cation \ce{[emim]^+} (see Fig.~\ref{fig:atoms_id}), as sampled in NPT-MD simulations of 250 ion pairs using each model for \ce{[emim][B(CN)_4]}.}
\label{fig:dihedrals_emim}
\end{figure*}

@@

A study of conformations was also carried out for \ce{[emim][NTf_2]} using the model of K\"{o}ddermann \textit{et al}. \cite{Koddermann_2007}.
A CDF obtained for the dihedral angles of sequences C1-N2-C5-C6 and C2-N2-C5-C6 in \ce{[emim]^+}, which is made available in the Supplementary Material, resembles the one in Fig.~\ref{fig:dihedrals_emim}(d).
The resemblance was expected because the intramolecular parameters employed for \ce{[emim]^+} are identical in these two cases, since the corresponding models are both based on the force field developed by Canongia-Lopes and P\'{a}dua \cite{Canongia_Lopes_2006}.
However, the most likely value of $\psi_1$ is shifted to $0^{\circ}$ in the case of \ce{[emim][NTf_2]}.
This highlights the fact that altering atomic charges can induce conformational changes and should, therefore, be accompanied by a revision of torsion parameters in some cases.
With regard to the anion \ce{[NTf_2]^-}, Deetlefs \textit{et al}. \cite{Deetlefs_2006} experimentally determined the conformer population of the liquid-phase and observed that \ce{[NTf_2]^-} is mostly found in its \textit{trans} conformation.
In Fig.~\ref{fig:die_ntf2}, we present the CDF obtained for the contiguous sequences C1-S1-N1-S2 and S1-N1-S2-C2 in \ce{[NTf_2]^-}, which qualitatively confirms the observation of Deetlefs \textit{et al}. \cite{Deetlefs_2006}.
This CDF also makes clear that \ce{[NTf_2]^-} exhibits a considerable degree of flexibility.
Due to this observed behaviors of both ions, all simulations of \ce{[emim][NTf_2]} in the present work are performed considering flexible cations and anions, as well as DSF electrostatics.
This approach results in a mean density of $1.495 \pm 0.009 ~\mathrm{g/cm^3}$ at $T = 298.0~\mathrm{K}$ and $P = 1.0~\mathrm{bar}$.
The relative error of this result with respect to the experimental value obtained by Tokuda \textit{et al}. \cite{Tokuda_2005} is only 1.12\%.

\begin{figure}
\centering
\includegraphics[width=\linewidth]{Ludwig_anion}
\caption{Combined distribution function of the dihedral angles of the contiguous chains C1-S1-N1-S2 and S1-N1-S2-C2 of anion \ce{[NTf_2]^-}, obtained from NPT-MD simulations of 250 ion pairs of \ce{[emim][NTf_2]} using the model of K\"{o}ddermann \textit{et al}. \cite{Koddermann_2007}.}
\label{fig:die_ntf2}
\end{figure}

@@

\begin{table*}
	\centering
	\caption{Results of solvation free energies and Henry's law constants of \ce{CO_2} in \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]} at $T = 298.15~\mathrm{K}$ and $P = 1~\mathrm{atm}$. The experimental Henry constants of \ce{CO_2} in \ce{[emim][B(CN)_4]} and in \ce{[emim][NTf_2]} are, respectively, $38.9 \pm 0.03 ~\mathrm{atm}$ \cite{Mahurin_2010} and $39.0 \pm 0.1 ~\mathrm{atm}$ \cite{Finotello_2008}.}
	\begin{adjustbox}{max width=\textwidth}
		\begin{tabular}{ccccccc}  
			\hline\hline
			Model & $\tilde{\rho}_\text{sim}$ & $\Delta G_\text{vdW}$  & $\Delta G_\text{Coul}$  & $\Delta G_\text{sim}$ & $K_s$ & error \\
			& (mol/dm\textsuperscript{3}) & (kcal/mol) & (kcal/mol) &  (kcal/mol) & (atm)  & (\%) \\
			\hline
%			\multicolumn{7}{c}{\ce{[emim][B(CN)_4]}: $K_{s}^{\text{exp}} = 38.9 \pm 0.03 ~\mathrm{atm}$ \cite{Mahurin_2010}} \\
			\multicolumn{7}{c}{\ce{[emim][B(CN)_4]}} \\
			Koller \textit{et al}. \cite{Koller_2012} & $4.5925$ & $0.45 \pm 0.01$ & $-1.06 \pm 0.02$ & $-0.61 \pm 0.02$ & $40 \pm 1$ & $2.8$ \\
			Batista \textit{et al}. \cite{Batista_2015} & $4.3700$ & $0.378 \pm 0.008$ & $-1.24 \pm 0.01$  & $-0.86 \pm 0.01$ & $24.9 \pm 0.5$ & $38.9$ \\
			Liu \textit{et al}. \cite{Liu_2014} & $4.3815$ & $0.226 \pm 0.009$ & $-1.11 \pm 0.01$ & $-0.88 \pm 0.01$ & $24.3 \pm 0.5$ & $37.6$  \\
			Weber and Kirchner \cite{Weber_2016} & $4.5221$ & $0.478 \pm 0.009$ & $-1.23 \pm 0.02$ & $-0.75 \pm 0.02$ & $31 \pm 1$ & $20.3$  \\
			\hline
%			\multicolumn{7}{c}{\ce{[emim][NTf_2]}: $K_{s}^{\text{exp}} = 39.0 \pm 0.1 ~\mathrm{atm}$ \cite{Finotello_2008} }\\
			\multicolumn{7}{c}{\ce{[emim][NTf_2]}} \\
			K\"{o}ddermann \textit{et al}. \cite{Koddermann_2007} & $3.8205$ & $0.389 \pm 0.005$ & $-0.901 \pm 0.006$ & $-0.512 \pm 0.008$ & $39.4 \pm 0.5$  & $0.97$  \\
			\hline\hline
			\label{table:henry} 
		\end{tabular}
	\end{adjustbox}
\end{table*}

@@

Such as the model of Koller \textit{et al}. \cite{Koller_2012} does in the case of \ce{[emim][B(CN)_4]}, the one of K\"{o}ddermann \textit{et al}. \cite{Koddermann_2007} accurately predicts the Henry's law constant of \ce{CO_2} in \ce{[emim][NTf_2]}.
A comparison between the values of $\Delta G_\text{sim}$ obtained from these models, shown in Table \ref{table:henry}, suggests that the solvation of \ce{CO_2} takes place more favorably in \ce{[emim][B(CN)_4]} than in \ce{[emim][NTf_2]}.
The fact that $K_s$ has almost the same value in both cases is due to a compensation promoted by the distinct molar densities of these ionic liquids.
Spatial distribution functions (SDFs) \cite{Svishchev_1993} can help us visualizing how differently the solvent ions interact with the solute molecule.
By using TRAVIS \cite{Brehm_2011}, we have obtained SDFs of both cation and anion around a \ce{CO_2} molecule, sampled from MD trajectories of either \ce{CO_2}+\ce{[emim][B(CN)_4]} or \ce{CO_2}+\ce{[emim][NTf_2]} systems simulated at $T = 298.0~\text{K}$, $P = 1.0~\mathrm{bar}$, and $x_\mathrm{CO_2} = 0.1$.
The slightly larger \ce{CO_2} concentration considered here is meant to improve statistics.
The obtained SDFs were depicted as isosurfaces using VMD \cite{HUMP96}, with isovalues determined so as to highlight the first solvation shell.
The resulting images are shown in Fig.~\ref{fig:sdf_ions}.
By comparing Figs.~\ref{fig:sdf_ions}(a) and \ref{fig:sdf_ions}(b), we note that the gap between the regions occupied by the anion and the cation is smaller in the case of \ce{[emim][B(CN)_4]}.
In addition, the first solvation shell corresponding to \ce{[B(CN)_4]^-} occurs at a slightly shorter distance than that corresponding to \ce{[NTf_2]^-}, as one can clearly see by a direct comparison done in Fig.~\ref{fig:sdf_ions}(c).
This is in part due to the difference in size of these two anions.
The result is a higher density of negative charges around the carbon atom of \ce{CO_2}, whose point-charge is positive as a consequence of the oxygen's high electronegativity.
This seems to induce a greater stability of \ce{CO_2} when solvated in \ce{[emim][B(CN)_4]}, rather than in \ce{[emim][NTf_2]}, possibly resulting in enhanced solubility.
However, the mentioned similarity in the Henry's law constants could cause such enhancement to go unnoticed.
This makes clear the importance of carrying out a study at high \ce{CO_2} concentrations.

@@

\begin{table}
	\centering
	\caption{Thermodynamic conditions employed in the solvation study at high concentrations of \ce{CO_2} in \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]}~\cite{Makino_2014,Schilderman_2007}.}
	\begin{tabular}{ c  c  c  }  
		\hline\hline
		IL & \ce{[emim][B(CN)_4]}  & \ce{[emim][NTf_2]} \\
		\hline
		model &  Ref.~\citenum{Koller_2012} & Ref.~\citenum{Koddermann_2007} \\
		T  & 313.15 K & 312.13 K \\
		P  & 21 atm & 37 atm \\
		$x_s$ & 0.352 & 0.479 \\
		$m_s$ & 2.4 mol/kg & 2.35 mol/kg \\
		\hline\hline
		\label{table:solv} 
	\end{tabular}
\end{table}

@@

\bibliography{ionic_liquids}
